\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-03-13 paper.qmd
Added approached followed for PBMC3k

@@

>>>>

### Approach followed

- First, referred to @Chen2023 paper which shows UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine.

- Observed that there is three well-separated clusters.

- Second, visualise the 15 PCs in high-dimensional space with our model and investigate whether can I see the three well-separated clusters or is it showing something different than what I saw in UMAP view suggested by the author.

- Learned from the model: Yes, it's different. There are three clusters but they are really close to each other. Also, in the largest cluster, observed non-linear continuity structure, which didn't captured by the author's UMAP representation.

- Learned from the data: Visualise with detour and colored the clusters in high-D. Identified more than three clusters which is suprising.

- In that case, need to suggest a better UMAP representation than author's suggestion.

- To find a better UMAP representation, I need to investigate different parameter choices for UMAP (or need to try another NLDR technique, but here my objective is to suggest the best parameter choice with UMAP).

- Therefore, I fitted our model for different UMAP parameter choices, record the error, and selected the parameter choice that gives the lowest error which is n:neighbor:5, min_dist:0.99, metric:cosine.

- Then, construct our model for this UMAP data (n:neighbor:5, min_dist:0.99, metric:cosine).

- In this representation, you can see that the clusters are closer, and the continuity of clusters are there.

- Also, what missed from looking at high-D data with detour, in this UMAP representation I can see more than three clusters.

- Then, need to check how correct am I identify the clusters in 2D and is that matched that I identify in high-D.

- Generate a confusion matrix.

@@

<!--obtain 2D and high-D distances-->

<!---
```{r}
#| warning: false
#| echo: false
#| message: false
#| eval: false

## Obtian 2D distances
proxy::dist(umap_pbmc[, -3]) |> 
as.data.frame()

## Obtain high-D distances
proxy::dist(training_data_pbmc[, -16]) |> 
as.data.frame()
```
--->

### 2024-03-13 paper.tex
Added approached followed for PBMC3k

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

@@

\hypertarget{approach-followed}{%
\subsection{Approach followed}\label{approach-followed}}

\begin{itemize}
\item
  First, referred to \citet{Chen2023} paper which shows UMAP
  representation of PBMC3k data with 15 PCs and parameter choice of
  n:neighbor:30, min\_dist:0.1, metric:cosine.
\item
  Observed that there is three well-separated clusters.
\item
  Second, visualise the 15 PCs in high-dimensional space with our model
  and investigate whether can I see the three well-separated clusters or
  is it showing something different than what I saw in UMAP view
  suggested by the author.
\item
  Learned from the model: Yes, it's different. There are three clusters
  but they are really close to each other. Also, in the largest cluster,
  observed non-linear continuity structure, which didn't captured by the
  author's UMAP representation.
\item
  Learned from the data: Visualise with detour and colored the clusters
  in high-D. Identified more than three clusters which is suprising.
\item
  In that case, need to suggest a better UMAP representation than
  author's suggestion.
\item
  To find a better UMAP representation, I need to investigate different
  parameter choices for UMAP (or need to try another NLDR technique, but
  here my objective is to suggest the best parameter choice with UMAP).
\item
  Therefore, I fitted our model for different UMAP parameter choices,
  record the error, and selected the parameter choice that gives the
  lowest error which is n:neighbor:5, min\_dist:0.99, metric:cosine.
\item
  Then, construct our model for this UMAP data (n:neighbor:5,
  min\_dist:0.99, metric:cosine).
\item
  In this representation, you can see that the clusters are closer, and
  the continuity of clusters are there.
\item
  Also, what missed from looking at high-D data with detour, in this
  UMAP representation I can see more than three clusters.
\item
  Then, need to check how correct am I identify the clusters in 2D and
  is that matched that I identify in high-D.
\item
  Generate a confusion matrix.
\end{itemize}
